### 2020-06-05 reviews/response.txt

We thank the reviewers for their exceptionally helpful and thorough
comments. Thorough reviews take time, interest, and expertise! Here,
we respond to reviewers' "Comments for Revision", marked by (C), and
due to limited space explain only a selection of detailed comments,
marked by (D). However, we have addressed all of the reviewers
comments in our revised paper either by making the requested change or
by providing additional explanation.

Our revision includes two substantive additions and numerous
improvements to explanations, graphics, and presentation. The first
substantive addition is a corroboration of our airflow implications on
temperature with temperature data from another publication. The second
substantive addition is a new result that job scheduling and folded
torus interconnect were contributing factors to temperature patterns,
and explain a peculiar interleaving column hazard variability.

Reviewer 1:

(C) Figure 2 is improved and corrected. The inventory dates are the same
for old and new batches, except that new units begin to appear in 2016 and
this is evident from the year colors.

Figure 7 is now split between old and new batches.

We now have location-specific information for average temperature
across cages and it agrees with our airflow effect on temperature
interpretation (V.-6th paragraph).

(D)

Reviewer 2:

(C) We have strengthened our explanations

### 2020-06-05 intro.tex
corrosion sentence in intro

Finding the root cause for the 2016 failures took a great deal of
testing and the involvement of materials science and microscopy
researchers. The failures were traced to a resistor on the GPU circuit
board (not the GPU chip iself) \fix{Was anything published about this?
  If not, let's use \cite{Vargas2009TheCO}}. Throughout this paper,
  when we refer to replacing the GPU, we refer to the entire
  circuitboard along with its GPU chip.

Our analysis focuses on GPU boards that were installed in the second
rework cycle and later, essentially units installed near the beginning
of 2014 and later. Figure~\ref{fig:chronology} is the only one
involving data prior to 2014. All analyses and figures that follow use
data from 2014 onward. We make the data as well as our analysis R
codes and Python codes to reproduce the graphics in this paper
publicly available (see, \cite[repository to be added in final
version]{repositoryreference,rcodes,pycodes}).  While the data is only
a 7 MB file, due to the size of Titan it represents over 100,000
collective years of GPU operation and may be the largest publicly
available GPU reliability data set. Our hope is that others will use
it for reliability and survival analysis research and go beyond the
results presented here.

>>>>

Finding the root cause for the 2016 failures took a great deal of
testing and the involvement of materials science and microscopy
researchers. The failures were traced to a resistor on the GPU circuit
board (not the GPU chip iself) due to silver sulphide
corrosion. \fix{Was anything published about this?  If not, let's use:
  Growth of such corrosion products in ambient air on microelectroncs
  parts is described in \cite{Vargas2009TheCO} and is cosistent with
  failures starting only after a critical amount of corrosion builds
  up, a situation that matches the experience on Titan}.

Throughout this paper, when we refer to replacing the GPU, we refer to
the entire circuitboard along with its GPU chip. Our analysis focuses
on GPU boards that were installed in the second rework cycle and
later, essentially units installed near the beginning of 2014 and
later. Figure~\ref{fig:chronology} is the only one involving data
prior to 2014. All analyses and figures that follow use data from 2014
onward. We make the data as well as our analysis R codes and Python
codes to reproduce the graphics in this paper publicly available (see,
\cite[repository to be added in final
version]{repositoryreference,rcodes,pycodes}).  While the data is only
a 7 MB file, due to the size of Titan it represents over 100,000
collective years of GPU operation and may be the largest publicly
available GPU reliability data set. Our hope is that others will use
it for reliability and survival analysis research and go beyond the
results presented here.

### 2020-06-05 survival.tex
corrosion sentence in intro

It appears that transport of cooling air provides a complete
explanation for relative differences in cage and node survival rates
in the old batch.  We reach this conclusion with reference to
\rev{Fig.}~\ref{fig:layout} and relative positions of cages and nodes
within a cabinet. Both cage and node differences in survival
probabilities can be explained by an inverse relationship with the
distance to the bottom of the cabinet, where cooling air is forced
through the cabinet to the top. The survival curves can be ordered
({\em cage 0, cage 1, cage 2}) in decreasing order of survival. As the
blades are placed vertically within the cabinet, pairs of nodes too
experience the same relationship with distance to the bottom, nodes 2
and 3 having lower failure rates than nodes 0 and 1. \rev{An
  increasing temperature gradient averaging about $4^\circ$C per
  cage is reported in \cite[p.7]{tiwari15Understand}, independently
  confirming our airflow temperature interpretation.}

>>>>

It appears that transport of cooling air provides a complete
explanation for relative differences in cage and node survival rates
in the old batch.  We reach this conclusion with reference to
\rev{Fig.}~\ref{fig:layout} and relative positions of cages and nodes
within a cabinet. Both cage and node differences in survival
probabilities can be explained by an inverse relationship with the
distance to the bottom of the cabinet, where cooling air is forced
through the cabinet to the top. The survival curves can be ordered
({\em cage 0, cage 1, cage 2}) in decreasing order of survival. As the
blades are placed vertically within the cabinet, pairs of nodes too
experience the same relationship with distance to the bottom, nodes 2
and 3 having lower failure rates than nodes 0 and 1.

\rev{An increasing average temperature gradient of about $4^\circ$C
  per cage is reported in \cite[Fig.10]{tiwari15Understand},
  independently confirming our airflow temperature
  interpretation. They also suggest that DBEs may be sensitive to
  temperature, although consider their evidence preliminary due to
  relatively low number of DBEs in 2014-1015 and high variability of
  $10^\circ$C to $15^\circ$C in their measured cage replicates.}
